Notes on Prediction Types

### 2012-02-08 docs/pert_notes/pert.tex
add SDC to the user's guide.  Define some section labels.
stub for Ryan's notes
some marginpars
Ryan's notes on toy_convect fun

@@

Predicting the full $\rho$ begins with Eq.~\ref{rho equation}:
\begin{equation}
\frac{\partial\rho}{\partial t} 
= -\Ub\cdot\nabla\rho \, \underbrace{- \rho\nabla\cdot\Ub}_{``\rho~\text{force''}} \, . \label{rho equation labeled}
\end{equation}
Using this, $\rho$ is predicted to the edges using {\tt
  make\_edge\_scal} in {\tt density\_advance}, with the underbraced
force computed in {\tt modify\_scal\_force} with {\tt fullform =
  .true.}.

@@

@@

@@

%% \begin{table*}[h]
%% \begin{center}
%% \caption{When predicting temp ... (conversion done by {\tt makeRhoHfromT}) \newline}
%% \begin{tabular}{c|c|c|c}
%% \hline
%% \hline
%% {type} & {Inputs } & {EOS inputs} & {output} \\
%% \hline
%% 3 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ &  $(\rho h)^\prime = \rho h - (\rho h)_0$ \\
%%    &                   & $   X = $ given & \\
%%    &                   &                 & \\
%%  4 & $X$ and $\rho^\prime$ & $\rho = \rho_0 + \rho^\prime$ & $h$ \\
%%    &                   & $   X = $ given & \\
%%    &                   &                 & \\
%% \hline
%% \end{tabular}
%% \end{center}
%% \end{table*}

%% \begin{table*}[h]
%% \begin{center}
%% \caption{Quantity that is created in {\tt mkflux} and passed into {\tt update\_scal}\newline\
%% }
%% \begin{tabular}{c|c|c|c|c}
%%        &            &              &\multicolumn{2}{c}{Forcing Term} \\
%% {type} & {Species } & {$(\rho h)$} & {$(\rho X)$} & {$(\rho h)$} \\
%% \hline 
%% 1 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
%% 2 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
%% 3 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0)\left [(\rho h)_0 + (\rho h)^\prime\right]$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}$ \\[2mm]
%% 4 & $(\Ubt+w_0)(\rho_0+\rho^\prime) X          $ & $(\Ubt+w_0) (\rho_0+\rho^\prime) h$ & 0 & $ \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r} $ \\[2mm]
%% \hline
%% \end{tabular}
%% \end{center}
%% \end{table*}

%% \begin{table*}[h]
%% \begin{center}
%% \caption{Quantity that goes into and out of {\tt make\_edge\_scal} \newline}
%% \begin{tabular}{c|c}
%% \hline
%% \hline
%% {h pred type} & {h/T } \\
%% \hline \\[-3mm]
%% 1 & $(\rho h)^\prime$ \\
%% 2 & $h$  \\
%% 3 & $T$  \\
%% 4 & $T$  \\
%% \hline
%% \end{tabular}
%% \end{center}
%% \end{table*}

%% \begin{table*}
%% \begin{center}
%% \caption{Quantity that goes into {\tt mkflux} on edges \newline}
%% \begin{tabular}{c|c|c}
%% \hline
%% \hline
%% {h pred type} & {Species } & {h/T } \\
%% \hline
%% 1 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
%% 2 & $X$ and $\rho^\prime$ & $h$ \\
%% 3 & $X$ and $\rho^\prime$ & $(\rho h)^\prime$ \\
%% 4 & $X$ and $\rho^\prime$ & $h$ \\
%% \hline
%% \end{tabular}
%% \end{center}
%% \end{table*}

>>>>

\label{sec:pred:density}

@@

Predicting the full $\rho$ begins with Eq.~\ref{rho equation}:
\begin{equation}
\frac{\partial\rho}{\partial t} 
= -\Ub\cdot\nabla\rho \, \underbrace{- \rho\nabla\cdot\Ub}_{``\rho~\text{force''}} \, . \label{rho equation labeled}
\end{equation}
Using this, $\rho$ is predicted to the edges using {\tt
  make\_edge\_scal} in {\tt density\_advance}, with the underbraced
force computed in {\tt modify\_scal\_force} with {\tt fullform =
  .true.}.  \MarginPar{we need to switch this over to doing the conservative prediction}

@@

\label{sec:pred:enthalpy}

@@

For {\tt species\_pred\_types}: {\tt predict\_rhoprime\_and\_X} or
{\tt predict\_rhoprime\_and\_rhoX}, we wish to construct $(\rho_0 +
\rho'_\mathrm{edge} ) h_\mathrm{edge}$. \MarginPar{alternately, we could construct $\rho_\mathrm{edge} = \sum_k (\rho X_k)_\mathrm{edge}$ for s\_p\_t = 2}

@@

%-----------------------------------------------------------------------------
% toy_convect
%-----------------------------------------------------------------------------
\section{Experience from {\tt toy\_convect}}

\label{sec:toyconvect}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Why is {\tt toy\_convect} Interesting?}

The {\tt toy\_convect} problem consists of a carbon-oxygen white dwarf with 
an accreted layer of solar composition. There is a steep composition 
gradient between the white dwarf and the accreted layer. The convection 
that begins as a result of the accretion is extremely sensitive to the 
amount of mixing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Initial Observations}

With {\tt use\_tfromp = T} and {\tt cflfac = 0.7} there is a large difference 
between {\tt species\_pred\_type = 1} and {\tt species\_pred\_type = 2,3} as 
seen in Figure \ref{fig:spec1_vs_23}. {\tt species\_pred\_type = 1} shows 
quick heating (peak T vs. t) and there is ok agreement between {\tt tfromh} 
and {\tt tfromp}. {\tt species\_pred\_type = 2,3} show cooling (peak T vs. t) 
and {\tt tfromh} looks completely unphysical (see Figure 
\ref{fig:tfromh_unphysical}). There are also strange filament type features in 
the momentum plots shown in Figure \ref{fig:mom_filaments}.

% compare spec 1 with spec 2,3
%-----------------------------------------
\begin{figure}[!h]
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/spec1_vs_23}
\caption{Compare {\tt species\_pred\_type = 1,2,3} with {\tt use\_tfromp = 
T, enthalpy\_pred\_type = 1, cflfac = 0.7}}
\label{fig:spec1_vs_23}
\end{figure}

% tfromh is unphysical
%-----------------------------------------
\begin{figure}[!h]
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/tfromh_unphysical}
\caption{{\tt tfromh} is unphysical when using {\tt species\_pred\_type = 
2,3, enthalpy\_pred\_type = 1, cflfac = 0.7, use\_tfromp = T}. Shown above 
is {\tt species\_pred\_type = 2}}
\label{fig:tfromh_unphysical}
\end{minipage}
\hspace{0.5cm}
% momentum filaments
%-----------------------------------------
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/mom_filaments}
\caption{There are strange filament type features at the bottom of the 
domain. {\tt species\_pred\_type = 2, enthalpy\_pred\_type = 1, cflfac = 0.7, 
use\_tfromp = T}}
\label{fig:mom_filaments}
\end{minipage}
\end{figure}

Using {\tt use\_tfromp = F} and {\tt dpdt\_factor $>$ 0} results in many runs 
crashing very quickly and gives unphyiscal temperature profiles as seen in 
Figure \ref{fig:tfrompF_unphys}.

% unphysical temp profiles with use_tfromp = F dpdt > 0
%-----------------------------------------
\begin{figure}[!h]
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/tfrompF_unphys}
\caption{Unphysical temperature profile with {\tt use\_tfromp = F} and 
{\tt dpdt\_factor = 0.1}. {\tt dpdt\_factor = 0.2,0.3} lead to the code 
crashing.}
\label{fig:tfrompF_unphys}
\end{minipage}
\hspace{0.5cm}
% temp vs time of castro ppm 0,1 and maestro spec = 2
%----------------------------------------------
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/tfrompF_cfl_1vs7}
\caption{Compare {\tt cflfac = 0.1} with {\tt cflfac = 0.7} for 
{\tt use\_tfromp = F, dpdt\_factor = 0.0, species\_pred\_type = 2, 
enthalpy\_pred\_type = 4}}
\label{fig:tfrompF_cfl_1vs7}
\end{minipage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Change {\tt cflfac} and {\tt enthalpy\_pred\_type}}

With {\tt species\_pred\_type = 1} and {\tt cflfac = 0.1}, 
there is much less heating (peak T vs. t) than the {\tt cflfac = 0.7} 
(default). There is also a lower overall Mach number (see Figure 
\ref{fig:spec1_cfl_1vs7}) with the {\tt cflfac = 0.1} and excellent agreement 
between {\tt tfromh} and {\tt tfromp}.

{\tt use\_tfromp = F, dpdt\_factor = 0.0, enthalpy\_pred\_type = 3,4 and 
species\_pred\_type = 2,3} shows cooling (as seen in {\tt use\_tfromp = T}) 
with a comparable rate of cooling (see Figure \ref{fig:compare_tfromp}) 
to the {\tt use\_tfromp = T} case. The 
largest difference between the two runs is that the {\tt use\_tfromp = F} 
case shows excellent agreement between {\tt tfromh} and {\tt tfromp} with 
{\tt cflfac = 0.7}. The filaments in the momentum plot of Figure 
\ref{fig:mom_filaments} are still present.

% Mach number of cfl 0.1 vs cfl 0.7
%-----------------------------------------
\begin{figure}[!h]
\begin{minipage}[!b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/spec1_cfl_1vs7}
\caption{Comparing the Mach number of {\tt cflfac = 0.1} and 
{\tt cflfac = 0.7}. {\tt species\_pred\_type = 1, enthalpy\_pred\_type = 1}}
\label{fig:spec1_cfl_1vs7}
\end{minipage}
\hspace{0.5cm}
% comparable cooling rates between tfromp = T and tfromp = F
%-----------------------------------------
\begin{minipage}[!b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/compare_tfromp}
\caption{Illustrate the comparable cooling rates between 
{\tt use\_tfromp = T} and {\tt use\_tfromp = F with dpdt\_factor = 0.0} 
using {\tt species\_pred\_type = 2, enthalpy\_pred\_type = 3,1}}
\label{fig:compare_tfromp}
\end{minipage}
\end{figure}

For a given {\tt enthalpy\_pred\_type} and {\tt use\_tfromp = F}, 
{\tt species\_pred\_type = 2} has a lower Mach number (vs. t) compared to 
{\tt species\_pred\_type = 3}.

Any {\tt species\_pred\_type} with {\tt use\_tfromp = F, dpdt\_factor = 0.0} 
and {\tt enthalpy\_pred\_type = 1} shows significant heating, although 
the onset of the heating is delayed in {\tt species\_pred\_type = 2,3} (see 
Figure \ref{fig:compare_tF_d0_h1_s123}). Only 
{\tt species\_pred\_type = 1} gives good agreement between {\tt tfromh} and 
{\tt tfromp}.

Comparing {\tt cflfac = 0.7} and {\tt cflfac = 0.1} with 
{\tt use\_tfromp = F, dpdt\_factor = 0.0, species\_pred\_type = 2} and 
{\tt enthalpy\_pred\_type = 4} shows good agreement overall (see Figure 
\ref{fig:tfrompF_cfl_1vs7}).

% compare tfromp=F%, dpdt=0, h=1, spec=1,2,3
%-----------------------------------------
\begin{figure}[!h]
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/compare_tF_d0_h1_s123}
\caption{Compare various {\tt species\_pred\_type} with {\tt use\_tfromp = F, 
dpdt\_factor = 0.0, enthalpy\_pred\_type = 1}}
\label{fig:compare_tF_d0_h1_s123}
\end{minipage}
\hspace{0.5cm}
% compare tfromp=F,dpdt=0,spec=2,h=4, cfl=0.1,0.7
%-----------------------------------------
\begin{minipage}[b]{0.5\linewidth}
\vspace{0pt}
\centering
\includegraphics[height=0.3\textheight]{\pertfigpath/compare_castro}
\caption{Compare the {\tt castro.ppm\_type CASTRO} runs with the 
{\tt species\_pred\_type MAESTRO} runs.}
\label{fig:compare_castro}
\end{minipage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additional Runs}

\subsubsection{{\tt bds\_type = 1}}

Using {\tt bds\_type = 1, use\_tfromp = F, dpdt\_factor = 0.0, 
species\_pred\_type = 2, enthalpy\_pred\_type = 4} and 
{\tt cflfac = 0.7} seems to cool off faster, perhaps due to less mixing. 
There is also no momentum filaments in the bottom of the domain.

\subsubsection{{\tt evolve\_base\_state = F}}

Using {\tt evolve\_base\_state = F, use\_tfromp = F, dpdt\_factor = 0.0, 
species\_pred\_type = 2} and {\tt enthalpy\_pred\_type = 4} seems to agree 
well with the normal {\tt evolve\_base\_state = T} run.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\tt toy\_convect} in {\tt CASTRO}}

{\tt toy\_convect} was also run using {\tt CASTRO} with 
{\tt castro.ppm\_type = 0,1}. These runs show temperatures that cool 
off rather than increase (see Figure \ref{fig:compare_castro}) which 
suggests using {\tt species\_pred\_type = 2,3} instead of 
{\tt species\_pred\_type = 1}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Recommendations}

All of these runs suggest that running under {\tt species\_pred\_type = 
2 or 3, enthalpy\_pred\_type = 3 or 4} with either {{\tt use\_tfromp = F} and 
{\tt dpdt\_factor = 0.0}} or {\tt use\_tfromp = T} gives the most 
consistent results.
